---
title: "Solution Description - Part (3)"
author: "Team A - Mufan Li, Mengye Ren, Tian Xia"
date: "March 19, 2016"
output: pdf_document
---

We first display the four time series plots.

```{r, echo=FALSE, warning=FALSE, include=FALSE, cache=FALSE}
# preprocess
library(ggplot2)
library(xtable)
options(xtable.comment = FALSE)
library(moments)
library(PerformanceAnalytics)
setwd("~/GitHub/quantathon-2016/")
file_name = "data_part3.team_A.csv"
data_part1 = read.csv(file_name)
data_part1$date = as.Date(as.character(data_part1$yyyymmdd),
                              format = "%Y%m%d")
names(data_part1) = gsub("return","returns",names(data_part1))
data_part1 = data_part1[data_part1$returns != 99,]

# temp_ret = as.numeric(rowMeans(in_data3[-1,] * W_df))
# mean(temp_ret)/sd(temp_ret)

in_data = read.csv("in_sample_data_headers.csv")
ind_df = in_data[-1:-2,grepl("IND_",names(in_data))]
W_df = data_part1[,grepl("Stock_",names(data_part1))]

data_part1[,grepl("Stock_",names(data_part1))] = W_df * (W_df * ind_df >= 0)

```

```{r, echo=FALSE, warning=FALSE}
# Daily LS Returns
ggplot(data_part1, aes(x=date, y=returns)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Returns") +
  ggtitle("1. Daily Long-Short Returns")
```

```{r, echo=FALSE, warning=FALSE}
# Cumulative LS Returns
# Find Equal Weighted Return
in_data2 = in_data[,grepl("SC",names(in_data))]

for (each_name in names(in_data2)) {
  in_data2[-1,each_name] =
    exp(diff(log(in_data2[,each_name])))-1
}
in_data3 = in_data2[-1,]
equal_weight_returns = as.numeric(rowMeans(in_data3))[-1]

ggplot(data_part1, aes(x=date)) + 
  geom_line(aes(y=cumsum(log(1+returns)),
                color="Long-Short Portfolio")) +
  geom_line(aes(y=cumsum(log(1+equal_weight_returns)),
                color="Equal-Weight Portfolio")) +
  scale_x_date("Date") + scale_y_continuous("Cumulative Log-Returns") +
  ggtitle("2. Cumulative Long-Short Returns - In Natural Logarithms") +
  guides(color=guide_legend(title=NULL)) + 
  theme(legend.position="bottom")
```

```{r, echo=FALSE, warning=FALSE}
# Mean Absolute Weights
mean_abs_weights = as.numeric(
    rowMeans(
      abs(
        data_part1[,grepl("Stock",names(data_part1))]
        )
      )
  )

ggplot(data_part1, aes(x=date, y=mean_abs_weights)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Mean Absolute Weights") +
  ggtitle("3. Daily Mean Absolute Weights")
```

```{r, echo=FALSE, warning=FALSE}
# Portfolio Direction
sum_weight = as.numeric(
    rowSums(data_part1[,grepl("Stock",names(data_part1))])
  )
sum_abs_weight = as.numeric(
  rowSums(abs(data_part1[,grepl("Stock",names(data_part1))]))
)
port_direction= sum_weight / sum_abs_weight
port_direction[is.nan(port_direction)] = 0

ggplot(data_part1, aes(x=date, y=port_direction)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Portfolio Direction") +
  ggtitle("4. Daily Portfolio Direction")
```

We can also look at the relevant statistics below.

```{r xtable, echo=FALSE, warning=FALSE, results = "asis"}
returns = data_part1$returns
log_returns = log(returns+1)
# returns = runif(nrow(data_part1),-0.1,0.1)
# log_returns = rnorm(length(data_part1$returns))

avg_return = mean(log_returns)
sd_return = sd(log_returns)
sharpe_ratio = avg_return / sd_return * sqrt(252)

skew_val = skewness(log_returns)
kurt_val = kurtosis(log_returns)

# DO NOT use log returns for drawdown!
# maxDrawdown(returns, geometric = T)
# findDrawdowns(edhec[,"Funds of Funds", drop=FALSE])
returns_df = data.frame(returns)
row.names(returns_df) = data_part1$date
tab_drawdowns = table.Drawdowns(returns_df,top=1)

days_drawdown = tab_drawdowns$`To Trough`
max_drawdown = tab_drawdowns$Depth

# mean(log(equal_weight_returns+1))/
#   sd(log(equal_weight_returns+1)) * sqrt(252)
# skewness(equal_weight_returns)
# kurtosis(equal_weight_returns)
cor_val = cor(equal_weight_returns,returns)

disp_table = data.frame(
  Names = c("Average Daily Log Returns",
            "Standard Deviation of Daily Log Returns",
            "Annualized Sharpe Ratio",
            "Skewness",
            "Kurtosis",
            "Maximum Drawdown - Number of Days",
            "Maximum Drawdown - Return",
            "Correlation with Equal Weighted Returns"),
  Values = c(avg_return,
            sd_return,
            sharpe_ratio,
            skew_val,
            kurt_val,
            days_drawdown,
            max_drawdown,
            cor_val)
)

# tab <- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2))
tab = xtable(disp_table,digits=4,
             caption="Summary Statistics - Using In-Sample Data")
print(tab)
```

This problem is clearly much more difficult
as we can observe the Sharpe ratio is significantly lower than part (2),
and in fact comparable to the data in part (1).
The maximum drawdown is also $175$ days and returns $-42\%$.
At the same time, if we observe plot 4, 
the strategy now tends to take more directional bets than in part (2).

The significant change due to a simple restriction
of trading directions is surprising, 
at least we were expecting a good in-sample result with 
possibly significant overfitting.

```{r, echo=FALSE, warning=FALSE, results = "asis"}
# Insert Table for Coefficients
coef_labels = paste0("a",1:12)
coef_data = as.numeric(read.csv("coeff_part3.team_A.csv"))
coef_df = data.frame(Parameter = coef_labels, Value = coef_data)
tab2 = xtable(coef_df,digits = 6,
              caption="Parameter Values - Using All In-Sample Data")
print(tab2)
```





















```{r, echo=FALSE, warning=FALSE, results = "asis"}
val_data = read.csv("data_part3_val.team_A.csv")
# write.csv(val_data,"data_part2_val.team_A.csv",row.names = F)
# nrow(val_data)

val_data$date = as.Date(as.character(val_data$yyyymmdd),
                              format = "%Y%m%d")
names(val_data) = gsub("return","returns",names(val_data))
val_data = val_data[val_data$returns != 99,]

returns = val_data$returns
log_returns = log(returns+1)

avg_return = mean(log_returns)
sd_return = sd(log_returns)
sharpe_ratio = avg_return / sd_return * sqrt(252)

skew_val = skewness(log_returns)
kurt_val = kurtosis(log_returns)

# DO NOT use log returns for drawdown!
# maxDrawdown(returns, geometric = T)
# findDrawdowns(edhec[,"Funds of Funds", drop=FALSE])
returns_df = data.frame(returns)
row.names(returns_df) = val_data$date
tab_drawdowns = table.Drawdowns(returns_df,top=1)

days_drawdown = tab_drawdowns$`To Trough`
max_drawdown = tab_drawdowns$Depth

# mean(log(equal_weight_returns+1))/
#   sd(log(equal_weight_returns+1)) * sqrt(252)
# skewness(equal_weight_returns)
# kurtosis(equal_weight_returns)
cor_val = cor(equal_weight_returns[1:length(returns)],returns)

disp_table = data.frame(
  Names = c("Average Daily Log Returns",
            "Standard Deviation of Daily Log Returns",
            "Annualized Sharpe Ratio",
            "Skewness",
            "Kurtosis",
            "Maximum Drawdown - Number of Days",
            "Maximum Drawdown - Return",
            "Correlation with Equal Weighted Returns"),
  Values = c(avg_return,
            sd_return,
            sharpe_ratio,
            skew_val,
            kurt_val,
            days_drawdown,
            max_drawdown,
            cor_val)
)

# tab <- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2))
tab = xtable(disp_table,digits=4,
             caption="Summary Statistics - Using Out-of-Sample Data")
print(tab)
```

At the same time, we also need to highlight a clear overfitting issue 
with this strategy. 
Similar to part (2), we also used the last $80\%$ of data 
as in-sample data and the first $20\%$ of data as out-of-sample.
As opposed to part (2), 
this time we observe a Sharpe ratio of $-0.74$.
The out-of-sample Sharpe ratio is likely 
made worse due to the choice of the period,
as we can observe both plot 2 and plot 5 of cumulative returns,
the choice of the first $20\%$ of data captures a period of poor performance.


```{r, echo=FALSE, warning=FALSE}
ggplot(val_data, aes(x=date)) + 
  geom_line(aes(y=cumsum(log(1+returns)),
                color="Long-Short Portfolio")) +
  geom_line(aes(y=cumsum(log(1+equal_weight_returns[1:length(returns)])),
                color="Equal-Weight Portfolio")) +
  scale_x_date("Date") + scale_y_continuous("Cumulative Log-Returns") +
  ggtitle("5. Cumulative Long-Short Returns - In Natural Logarithms - Out-of-Sample") +
  guides(color=guide_legend(title=NULL)) + 
  theme(legend.position="bottom")
```

```{r, echo=FALSE, warning=FALSE, results = "asis"}
# Insert Table for Coefficients
coef_labels = paste0("a",1:12)
coef_data = as.numeric(read.csv("coeff_part3_val.team_A.csv"))
coef_df = data.frame(Parameter = coef_labels, Value = coef_data)
tab2 = xtable(coef_df,digits = 6,
              caption="Parameter Values - Using 80\\% of In-Sample Data")
print(tab2)
```














