---
title: "Solution Description - Part (2)"
author: "Team A - Mufan Li, Mengye Ren, Tian Xia"
date: "March 19, 2016"
output: pdf_document
---

We first display the four time series plots.

```{r, echo=FALSE, warning=FALSE, include=FALSE, cache=FALSE}
# preprocess
library(ggplot2)
library(xtable)
options(xtable.comment = FALSE)
library(moments)
library(PerformanceAnalytics)
setwd("~/GitHub/quantathon-2016/")
file_name = "data_part2.team_A.csv"
data_part1 = read.csv(file_name)

data_part1$date = as.Date(as.character(data_part1$yyyymmdd),
                              format = "%Y%m%d")
names(data_part1) = gsub("return","returns",names(data_part1))
data_part1 = data_part1[data_part1$returns != 99,]

```

```{r, echo=FALSE, warning=FALSE}
# Daily LS Returns
ggplot(data_part1, aes(x=date, y=returns)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Returns") +
  ggtitle("1. Daily Long-Short Returns")
```

```{r, echo=FALSE, warning=FALSE}
# Cumulative LS Returns
# Find Equal Weighted Return
in_data = read.csv("in_sample_data_headers.csv")
in_data2 = in_data[,grepl("SC",names(in_data))]

for (each_name in names(in_data2)) {
  in_data2[-1,each_name] =
    exp(diff(log(in_data2[,each_name])))-1
}
in_data3 = in_data2[-1,]
equal_weight_returns = as.numeric(rowMeans(in_data3))[-1]

ggplot(data_part1, aes(x=date)) + 
  geom_line(aes(y=cumsum(log(1+returns)),
                color="Long-Short Portfolio")) +
  geom_line(aes(y=cumsum(log(1+equal_weight_returns)),
                color="Equal-Weight Portfolio")) +
  scale_x_date("Date") + scale_y_continuous("Cumulative Log-Returns") +
  ggtitle("2. Cumulative Long-Short Returns - In Natural Logarithms") +
  guides(color=guide_legend(title=NULL)) + 
  theme(legend.position="bottom")
```

```{r, echo=FALSE, warning=FALSE}
# Mean Absolute Weights
mean_abs_weights = as.numeric(
    rowMeans(
      abs(
        data_part1[,grepl("Stock",names(data_part1))]
        )
      )
  )

ggplot(data_part1, aes(x=date, y=mean_abs_weights)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Mean Absolute Weights") +
  ggtitle("3. Daily Mean Absolute Weights")
```

```{r, echo=FALSE, warning=FALSE}
# Portfolio Direction
sum_weight = as.numeric(
    rowSums(data_part1[,grepl("Stock",names(data_part1))])
  )
sum_abs_weight = as.numeric(
  rowSums(abs(data_part1[,grepl("Stock",names(data_part1))]))
)
port_direction= sum_weight / sum_abs_weight
port_direction[is.nan(port_direction)] = 0

ggplot(data_part1, aes(x=date, y=port_direction)) + geom_line() +
  scale_x_date("Date") + scale_y_continuous("Portfolio Direction") +
  ggtitle("4. Daily Portfolio Direction")
```

We can also look at the relevant statistics below.

```{r xtable, echo=FALSE, warning=FALSE, results = "asis"}
returns = data_part1$returns
log_returns = log(returns+1)
# returns = runif(nrow(data_part1),-0.1,0.1)
# log_returns = rnorm(length(data_part1$returns))

avg_return = mean(log_returns)
sd_return = sd(log_returns)
sharpe_ratio = avg_return / sd_return * sqrt(252)

skew_val = skewness(log_returns)
kurt_val = kurtosis(log_returns)

# DO NOT use log returns for drawdown!
# maxDrawdown(returns, geometric = T)
# findDrawdowns(edhec[,"Funds of Funds", drop=FALSE])
returns_df = data.frame(returns)
row.names(returns_df) = data_part1$date
tab_drawdowns = table.Drawdowns(returns_df,top=1)

days_drawdown = tab_drawdowns$`To Trough`
max_drawdown = tab_drawdowns$Depth

# mean(log(equal_weight_returns+1))/
#   sd(log(equal_weight_returns+1)) * sqrt(252)
# skewness(equal_weight_returns)
# kurtosis(equal_weight_returns)
cor_val = cor(equal_weight_returns,returns)

disp_table = data.frame(
  Names = c("Average Daily Log Returns",
            "Standard Deviation of Daily Log Returns",
            "Annualized Sharpe Ratio",
            "Skewness",
            "Kurtosis",
            "Maximum Drawdown - Number of Days",
            "Maximum Drawdown - Return",
            "Correlation with Equal Weighted Returns"),
  Values = c(avg_return,
            sd_return,
            sharpe_ratio,
            skew_val,
            kurt_val,
            days_drawdown,
            max_drawdown,
            cor_val)
)

# tab <- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2))
tab = xtable(disp_table,digits=4,
             caption="Summary Statistics - Using In-Sample Data")
print(tab)
```

To find the parameter values,
we tried both gradient descent and 
random line search algorithms in MATLAB.
Since we found multiple different local minima,
the problem is clearly non-convex;
however all of the Sharpe ratios are fairly similar,
so it is not a significant issue in this problem.

```{r xtable, echo=FALSE, warning=FALSE, results = "asis"}
# Insert Table for Coefficients
# coef_data = read.csv("coeff_part2.team_A.csv")
```

Judging from in-sample data, 
the strategy is performing quite well.
We first observe plot 2 of cumulative returns 
and the annualized Sharpe ratio, 
both of which are significantly outperforming 
the strategy in part (1) and the equal weight portfolio.
From a correlation of only $0.0113$
with the equal weight portfolio,
and observing plot 4
this strategy rarely has a portfolio direction of 
more than $0.2$,
we can see that the strategy is very market neutral.
We also note the maximum drawdown for this strategy is 
extremely low at $4$ days and $-7.3\%$.

The only negative sign we observe from in-sample statistics 
is the large skew and excess kurtosis,
implying extremely heavy tails.
However, but looking a histogram below (plot 5), 
we realize that there is data point returning $20\%$.
Removing that point, we would have a skewness of $0.556$
and an excess kurtosis of $7.640$

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# skewness(log_returns[log_returns<0.1])
# kurtosis(log_returns[log_returns<0.1])
ggplot(NULL,aes(x=log_returns)) + geom_histogram() +
  xlab("Daily Log Returns") + 
  scale_y_continuous("Frequency") +
  ggtitle("5. Histogram of Daily Log Returns")
```




















